The Weakly Nonlinear Magnetorotational Instability in a local geometry

### 2017-03-24 paper/thingap_ref_response/response_to_referee.tex

Since these two papers are strongly linked, I am submitting a single, combined report. Although I recommend extensive revisions, I consider this an impressive piece of work by a young researcher who should be encouraged to continue working on this problem.

The following remarks apply to both papers:

\begin{enumerate}				
\item One of the main ideas behind the “weakly nonlinear” approach is that near marginal stability there is only one or a very small band of unstable modes that self interact. There is a substantial body of work in the fusion plasma community that considers coupling to damped modes as an important ingredient in saturation. The names to look for are Hatch, Makwana, and Terry. I am not suggesting the authors redo their analysis, to include damped modes, but they should comment on this alternative, perhaps in the Introduction.

\textbf{Coupling to damped modes as explored in the plasma literature is not analytically possible in the walled TC flow, because the boundaries require numerical computation of the eigenvectors, precluding the simple identification of modes with branches of an algebraic dispersion relation. Such a program is beyond the scope of this work, though it is a very interesting possibility we had not previously considered. However, such inter-mode coupling \emph{can} be analyzed in the shearing box when ambipolar diffusion is considered, as described in the new section of the text.}

@@

\textbf{The amplitude $\alpha$ is bounded by the value +/- $(b/c)^{1/2}$, although the amplitude does not necessarily converge to a constant value, as shown in Figures 6 and 7. We note, however, that although the value of the quantity $(b/c)^{1/2}$ is heavily influenced by the boundary layers in the walled TC flow, this is specific to the wall-like boundary conditions and not inherent to the theory. We have added a paragraph at the end of section 4 to elaborate on this point, discuss what the $(b/c)^{1/2}$ saturation amplitude means physically, and describe how the derived amplitude applies to the shearing box MRI.}

\item Both papers are light on physical interpretation. This is especially an issue with Paper I -- see below. I would particularly like to see:

\begin{itemize}						 							
\item A more extended discussion of how angular momentum transport works. After all, this is the primary raison d’etre for most studies of the MRI.

\textbf{We have added a figure to the thin-gap paper showing the morphology of the average angular momentum transport $\dot{J}$. }

@@

\textbf{Yes, the saturated Bz is marginally stable, and we have added a sentence to that effect in the text. We cannot comment on dynamo action in either paper because of the symmetries in our setup. No dynamo can arise from an axisymmetric system (e.g. Cowling's anti-dynamo theorem).}

\end{itemize}

\item A small point – the parameter called $\beta$ is not the standard definition; why not introduce the Cowling number?

\textbf{We have made this change.}

\item Finally, would it be feasible to a compute a nonlinear solution with Dedalus (or some other platform) and see how accurate the perturbation theory really is, for different values of $\epsilon$? Although the authors compare their results with Ebrahimi 2009, it would be desirable to have their own simulations for comparison.

  \textbf{We have added a new section including a direct numerical simulation run with Dedalus, and made a preliminary comparison at modest Pm. }

\end{enumerate}

### 2017-03-24 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
referee response changes.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{thingap_stressfig_Pm_1E-2}
\caption{\edit1{Total (Reynolds + Maxwell) stress in the domain as predicted from the weakly nonlinear theory.}} \label{fig:wnl_stress}
\end{figure}

@@

>>>>

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{thingap_stressfig_Pm_1E-2}
\caption{\edit1{Total (Reynolds + Maxwell) stress in the domain as predicted from the weakly nonlinear theory at $\Pm = 10^{-2}$.}} \label{fig:wnl_stress}
\end{figure}

@@

\edit1{Figure~\ref{fig:wnl_stress} shows the total stress $u_x u_y - \Co B_x B_y$ for the $\Pm = 10^{-2}$ model. The stress shows signficantly more structure throughout the domain than the  variables $u_x$, $u_y$, $B_x$ and $B_y$ that comprise it, demonstrating that a non-trivial correlation exists even in the weakly non-linear state. As in simulations at higher $\reym$, figure~\ref{fig:sim_theory_energy_jdot} shows that the Maxwell stress dominates over the Reynolds stresses even though the kinetic energy significantly exceeds the magnetic energy.  }
